---
title: 'Topic 4: Sentiment Analysis II'
output:
  pdf_document:
    latex_engine: xelatex
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### IPCC Report Twitter

```{r packages, results='hide', message=FALSE, warning=FALSE}
library(quanteda)
#devtools::install_github("quanteda/quanteda.sentiment") #not available currently through CRAN
library(quanteda.sentiment)
library(quanteda.textstats)
library(tidyverse)
library(tidytext)
library(lubridate)
library(wordcloud) #visualization of common words in the data set
library(reshape2)
library(sentimentr) #for question 5
```

Last week we used the tidytext approach to sentiment analysis for Nexis Uni .pdf data on coverage of the recent IPCC report. This week we will look at the conversation on Twitter about the same report. We'll start with the familiar tidy approach, and then introduce the quanteda package later.

```{r tweet_data}

raw_tweets <- read.csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/dat/IPCC_tweets_April1-10_sample.csv", header=TRUE)

dat<- raw_tweets[,c(4,6)] # Extract Date and Title fields

tweets <- tibble(text = dat$Title,
                  id = seq(1:length(dat$Title)),
                 date = as.Date(dat$Date,'%m/%d/%y'))


head(tweets$text, n = 10)

#simple plot of tweets per day
tweets %>%
  count(date) %>%
  ggplot(aes(x = date, y = n))+
  geom_line()

```

*Question 1. Think about how to further clean a twitter data set. Let's assume that the mentions of twitter accounts is not useful to us. Remove them from the text field of the tweets tibble.*
```{r cleaning_tweets}

#let's clean up the URLs from the tweets
tweets$text <- gsub("http[^[:space:]]*", "",tweets$text)
tweets$text <- str_to_lower(tweets$text)

#Remove twitter account mentions from the text field of the tweets tibble.
tweets$text <- gsub("@[^[:space:]]*", "",tweets$text)

#load sentiment lexicons
bing_sent <- get_sentiments('bing')
nrc_sent <- get_sentiments('nrc')

#tokenize tweets to individual words
words <- tweets %>%
  select(id, date, text) %>%
  unnest_tokens(output = word, input = text, token = "words") %>%
  anti_join(stop_words, by = "word") %>%
  left_join(bing_sent, by = "word") %>%
  left_join(
    tribble(
      ~sentiment, ~sent_score,
      "positive", 1,
      "negative", -1),
    by = "sentiment")
```

```{r sentiment_calculations}
#take average sentiment score by tweet
tweets_sent <- tweets %>%
  left_join(
    words %>%
      group_by(id) %>%
      summarize(
        sent_score = mean(sent_score, na.rm = T)),
    by = "id")

neutral <- length(which(tweets_sent$sent_score == 0))
positive <- length(which(tweets_sent$sent_score > 0))
negative <- length(which(tweets_sent$sent_score < 0))

Sentiment <- c("Positive","Neutral","Negative")
Count <- c(positive,neutral,negative)
output <- data.frame(Sentiment,Count)
output$Sentiment<-factor(output$Sentiment,levels=Sentiment)
ggplot(output, aes(x=Sentiment,y=Count))+
  geom_bar(stat = "identity", aes(fill = Sentiment))+
  scale_fill_manual("legend", values = c("Positive" = "green", "Neutral" = "black", "Negative" = "red"))+
  ggtitle("Barplot of Sentiment in IPCC tweets")
```

```{r plot_sentiment_by_day}
# tally sentiment score per day
daily_sent <- tweets_sent %>%
  group_by(date) %>%
  summarize(sent_score = mean(sent_score, na.rm = T))

daily_sent %>%
  ggplot( aes(x = date, y = sent_score)) +
  geom_line() +
    labs(x = "Date",
    y = "Avg Sentiment Score",
    title = "Daily Tweet Sentiment",
    subtitle = "IPCC Tweets")

```

*Quetsion 3. Adjust the wordcloud in the "wordcloud" chunk by coloring the positive and negative words so they are identifiable.*

Now let's try a new type of text visualization: the wordcloud.

```{r wordcloud}
words %>%
   anti_join(stop_words) %>%
   count(word) %>%
   with(wordcloud(word, n, max.words = 100))


```

```{r wordcloud_comp}

words %>%
inner_join(get_sentiments("bing")) %>%
count(word, sentiment, sort = TRUE) %>%
acast(word ~ sentiment, value.var = "n", fill = 0) %>%
comparison.cloud(colors = c("red", "green"),
                   max.words = 100)
```


*Question 2.Compare the ten most common terms in the tweets per day.  Do you notice anything interesting?*

```{r}
#Word count 
text_wordcounts <- words  %>% count(word, sort = TRUE)

#Get only the ten most common terms 
text_wordcounts <- text_wordcounts[1:10, ]

#select terms in words df that match the 10 most common terms 
common_terms <- words %>% 
  subset(word %in% text_wordcounts$word)
  
# Group by date 
common_terms_date <- common_terms %>% 
  group_by(date) %>% 
  count(word)

#Plot of ten most common terms in the tweets per day
ggplot(data = common_terms_date, aes(x = date, y = n)) +
  geom_line(aes(color = word)) +
  labs(title = "Tweets Per Day of 10 Most Common Terms (2022)",
       x = "Date",
       y = "Number of Tweets With Common Terms") +
  labs(color = "Term")
```
The thing I notice is that all of the most common words sem to be pretty straight forward and I could have guessed them. Id be interested to see what they are if you take out "ipcc", "report", and other words like that

#### The quanteda package

quanteda is a package (actually a family of packages) full of tools for conducting text analysis. quanteda.sentiment (not yet on CRAN, download from github) is the quanteda modular package for conducting sentiment analysis.

quanteda has its own built in functions for cleaning text data. Let's take a look at some. First we have to clean the messy tweet data:

```{r create_corpus}
corpus <- corpus(dat$Title) #enter quanteda
summary(corpus)
```

```{r quanteda_cleaning}
tokens <- tokens(corpus) #tokenize the text so each doc (page, in this case) is a list of tokens (words)

#examine the uncleaned version
tokens

#clean it up
tokens <- tokens(tokens, remove_punct = TRUE,
                      remove_numbers = TRUE)

tokens <- tokens_select(tokens, stopwords('english'),selection='remove') #stopwords lexicon built in to quanteda

#tokens <- tokens_wordstem(tokens) #stem words down to their base form for comparisons across tense and quantity

tokens <- tokens_tolower(tokens)

```

We can use the kwic function (keywords-in-context) to briefly examine the context in which certain words or patterns appear.

```{r initial_analysis}
head(kwic(tokens, pattern = "climate", window = 3))

head(kwic(tokens, pattern = phrase("climate change"), window = 3))


```

*Question 4. Let's say we are interested in the most prominent entities in the Twitter discussion.  Which are the top 10 most tagged accounts in the data set? Hint: the "explore_hashtags" chunk is a good starting point.*


```{r explore_hashtags}
hash_tweets <- tokens(corpus, remove_punct = TRUE) %>% 
               tokens_keep(pattern = "#*")

dfm_hash<- dfm(hash_tweets)

tstat_freq <- textstat_frequency(dfm_hash, n = 100)
head(tstat_freq, 10)

#tidytext gives us tools to convert to tidy from non-tidy formats
hash_tib<- tidy(dfm_hash)

hash_tib %>%
   count(term) %>%
   with(wordcloud(term, n, max.words = 100))

#this is where the hw answer starts 
tagged_tweets <- tokens(corpus, remove_punct = TRUE) %>% 
               tokens_keep(pattern = "@*")
dfm_tagged<- dfm(tagged_tweets)
tstat_freq <- textstat_frequency(dfm_tagged, n = 100)
tstat_freq <- tstat_freq[1:10, ]
#tidytext gives us tools to convert to tidy from non-tidy formats
tagged_tib<- tidy(dfm_tagged)
tagged_tib %>%
   count(term) %>%
   with(wordcloud(term, n, max.words = 50))
```

```{r}
# Plot Top 10 Most Tagged Twitter Accounts in IPCC Dataset
ggplot(data = tstat_freq, aes(y = reorder(feature, -rank), x = frequency)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Most Tagged Twitter Accounts in IPCC Dataset",
       x = "Number of Tags",
       y = "Account") +
  theme_bw()
```


Create the sparse matrix representation known as the document-feature matrix. quanteda's textstat_polarity function has multiple ways to combine polarity to a single score. The sent_logit value to fun argument is the log of (pos/neg) counts.

```{r}

dfm <- dfm(tokens)

topfeatures(dfm, 12)

dfm.sentiment <- dfm_lookup(dfm, dictionary = data_dictionary_LSD2015)

head(textstat_polarity(tokens, data_dictionary_LSD2015, fun = sent_logit))


```

*Question 5. The Twitter data download comes with a variable called "Sentiment" that must be calculated by Brandwatch.  Use your own method to assign each tweet a polarity score (Positive, Negative, Neutral) and compare your classification to Brandwatch's (hint: you'll need to revisit the "raw_tweets" data frame).*

```{r sentiment score}
#data cleaning and wrangling 

data_text <- raw_tweets[,c(4,6)] 

tweets_text <- tibble(text = data_text$Title,
                  id = seq(1:length(data_text$Title)),
                 date = as.Date(data_text$Date,'%m/%d/%y'))

tweets_text$text <- gsub("http[^[:space:]]*", "",tweets_text$text)
tweets_text$text <- str_to_lower(tweets_text$text)

tweets_text$text <- gsub("@[^[:space:]]*", "",tweets_text$text)


mytext <- get_sentences(tweets_text$text) %>% 
  sentiment() %>% 
  rename(sentiment_score = sentiment) #uses the sentimentr package to get the sentiment based on polartiy, with teh next line

sentiment_assigned <- mytext %>% 
  group_by(element_id) %>% 
  summarise(sentiment_score = mean(sentiment_score))

sentiment_assigned$sentiment <- ifelse(sentiment_assigned$sentiment_score <0, 
                                       "Negative", 
                                       ifelse(sentiment_assigned$sentiment_score > 0, "Positive", "Neutral"))



#Now we have to get the number of tweets with each sentiment type 
raw_sentiment <- sentiment_assigned %>%
  group_by(sentiment) %>% 
  summarise(sentiment_count = n())

#Plot the count per sentiment 
ggplot(data = raw_sentiment, aes(x = Sentiment, y = sentiment_count)) +
         geom_bar(stat = "identity", aes(fill = Sentiment)) +
  labs(title = "Sentiment Classification",
         x= "Sentiment",
         y= "Count")
```


```{r}
#Find the number of tweets with each sentiment type for Brandwatch 
raw_sentiment <- raw_tweets %>%
  group_by(Sentiment) %>% 
  summarise(sentiment_count = n())
#Plot the count per sentiment 
ggplot(data = raw_sentiment, aes(x = Sentiment, y = sentiment_count)) +
         geom_bar(stat = "identity", aes(fill = Sentiment)) +
  labs(title = "Brandwatch Sentiment Classification",
         x= "Sentiment",
         y= "Count")
```
By doing it like this, and using the sentimentr package it is clear to see that there are more words assigned to either negative or positive when compared to the Brandwatch dataset. The overall trend is still the same, where its slightly more negative, however, there is a larger number of overall words in the negative and positive categories.



